Dike experiment

### 2023-01-25 manuscript.Rmd
Landscape orientation with officedown

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  bookdown::word_document2:
    reference_docx: StylesTemplate.docx
    global_numbering: true
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

@@

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:map)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table \@ref(tab:flood-weather), @lfu2021waterlevel;@lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland [Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016]. The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <!--# check numbers  -->, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits)) . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(tab:substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ) <!--# include superscript -->. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species <!--# Refer to Table Sx -->.

<!--# We measured FCS, recovery, persistence -->

@@

A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time. Rare species with an accumulated cover over all plots of less than 0.5% were excluded (2D-stress: 0.21).

@@

# Open research {open-research, .unnumbered}

Data and code is available on Zenodo: <!--# https://doi.org/10.XXXX -->

@@

```{r table-traits, tab.id = "traits", tab.cap = "Traits"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  autofit()
```

@@

```{r table-substrate, tab.id = "substrate", tab.cap = "Substrate"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  autofit()
```

@@

```{r figure-nmds, fig.cap = "NMDS"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

```{r table-establishment, fig.cap = "table-s2"}
knitr::include_graphics(here::here("outputs", "tables", "table_a2_establishment.html"))
```

>>>>

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  officedown::rdocx_document:
    reference_docx: StylesTemplate.docx
  #bookdown::word_document2:
    #global_numbering: true
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
---

@@

-   <!--# We measured FCS, recovery, persistence -->

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)); 314 m above sea level; WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and an annual precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@hari2020; @cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table \@ref(tab:flood-weather), @lfu2021waterlevel; @lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland [Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016]. The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species , respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits)) . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(tab:substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g^-2^). The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species (Appendix Table \@ref(fig:table-establishment)).

## External data

Reference vegetation surveys were extracted from sPlotOpen [@Sabatini.2021] and our owen surveys in the surroundings on Danube dikes [@bauer2023]. We chose

@@

A non-metric multidimensional scaling ordination (NMDS) with Jaccard dissimilarity was used to visualize variation in species composition in space and time. Rare species with an accumulated cover over all plots of less than 0.5% were excluded (2D-stress: 0.21). <!--# Refer to Table Sx -->

@@

# Open research {#sec-open-research-open-research-.unnumbered .unnumbered}

@@

```{r table-traits, tab.id = "traits", tab.cap = "Traits"}
#read_csv(here::here("data", "raw", "data_raw_table_3.csv")) %>%
  #flextable() %>%
  #bold(part = "header") %>%
  #autofit()
```

@@

```{r table-substrate, tab.id = "substrate", tab.cap = "Substrate"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  qflextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  autofit()
```

@@

```{r figure-nmds, fig.cap = "NMDS"}
knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

```{r table-establishment, fig.cap = "table-s2"}
#knitr::include_graphics(here::here("outputs", "tables", "table_a2_establishment.html"))
```
